---
title: "Penguins Flipper Length ANOVA"
author: "Jason Johns"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r, include = FALSE}

library(tidyverse)
library(car)
library(onewaytests)


penguins <- read_csv("penguins.csv")

```

Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelle Penguins near Palmer Archipelago?

Maybe I can use a 1-way ANOVA...
- Independent obs.
- Equal variance (Levene's test)
- Normally distributed data (histograms/qq plots)

```{r}

# Shows us the number of observations of each sex for each species
counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

```

Exploratory graphs and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

hists

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

qqs

# Based on qq plots & histograms, observations for each group appear normally distributed. Also, n > 30 for all groups, so means will be normally distributed by CLT.

```

Evaluate variances and perform Levene's Test for equal variances

```{r}
# If the largest sample variances is <4x greater than the smallest sample variance, then these are "close enough" to consider variances equal.

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances # The largest variance (33.1) is definitely not >4x greater than the smallest variance (15.2).

# Levene's Test
#leveneTest function in 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

# H0: no difference in variances (variances are equal)
# HA: variances are NOT equal

# There is no significant difference in variances across species.

```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}

# H0: Means across all groups are equal
# HA: at least 2 group means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# Post-hoc testing by Tukey's HSD
# H0: No significant difference in pairwise means
# HA: There is a significant difference in pairwise means 

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# aov.test function in the 'onewaytests' package DOES allow for inline references

```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way ANOVA, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).